---
title: "test_SBC"
format: html
---

## Testing SBC
### On the iid normal example

Load our package:

```{r}
source("R/sbc.R")
```


Taken directly from `tutotial_bayes_gibbs_sbc`:

```{r}
#| label: other-samplers

prior_sampler <- function(prior_hyper_params){
  
  a0 <- prior_hyper_params[1]
  b0 <- prior_hyper_params[2]
  mu0 <- prior_hyper_params[3]
  v20 <- prior_hyper_params[4]
  
  # simulate mu ~ N(mu0, v20)
  mu <- rnorm(1, mu0, sqrt(v20))
  
  # independently simulate sigma2 ~ IG(a0, b0)
  sigma2 <- 1 / rgamma(1, a0, rate = b0)

  return(c(mu = mu, sigmasq = sigma2))
}

likelihood_sampler <- function(n, theta){
  mu <- theta[1]
  sigma2 <- theta[2]
  
  y <- rnorm(n, mu, sqrt(sigma2))
  
  return(y)
}
```

And the posterior sampler:

```{r}
#| label: posterior-sampler

posterior_sampler <- function(ndraws, 
                              y, 
                              prior_hyper_params, 
                              init = c(mean(y), var(y))){

  n <- length(y)
  sum.y <- sum(y)
  a0 <- prior_hyper_params[1]
  b0 <- prior_hyper_params[2]
  mu0 <- prior_hyper_params[3]
  v20 <- prior_hyper_params[4]

  # preallocate storage
  THETA <- matrix(0, nrow = ndraws, ncol = length(init))
  
  # initialize
  theta <- init

  for(s in 1:ndraws){
    
    # draw mean given variance and data
    
    v2n = 1 / (1/v20 + n/theta[2]) 
    mun = v2n * (mu0/v20 + sum.y/theta[2])
    
    theta[1] <- rnorm(1, mun, sqrt(v2n))
    
    # draw variance given mean and data
    
    an = a0 + n/2
    bn = b0 + sum((y - theta[1])^2) / 2
    
    theta[2] <- 1 / rgamma(1, shape = an, rate = bn)
    
    # store the latest draw in the sth row of the THETA matrix
    
    THETA[s, ] <- theta
    
  }
  
  return(THETA)
}
```

Set hyperparameters:

```{r}
prior_hyper_params <- c(a0 = 1, b0 = 1, mu0 = 0, v20 = 1)
```

Run it:

```{r}
res <- run_sbc(
  prior_sampler      = prior_sampler,
  likelihood_sampler = likelihood_sampler,
  posterior_sampler  = posterior_sampler,
  n_sims             = 1000,        # small for testing
  n_obs              = 50,
  n_draws            = 1000,        # small for testing
  prior_hyper_params = prior_hyper_params,
  parallelize        = FALSE
)
```
```{r}
SBC::plot_rank_hist(res$sbc_result)
```
```{r}
SBC::plot_ecdf_diff(res$sbc_result)
```
Ok great! Time to see if we can now use our package to run some test functions on it.

```{r}
source("R/testfunctions.R")
```

We need to define our test functions in a specific way. Like this:

```{r}
test_fns <- make_test_functions(
  mu_sq = function(theta, y) theta["mu"]^2,
  product = function(theta, y) theta["mu"] * theta["sigmasq"],
  log_like = function(theta, y) sum(dnorm(y, 
                                          mean = theta["mu"], 
                                          sd = sqrt(theta["sigmasq"]),
                                          log = TRUE))
)
```

Recompute ranks using our new function:

```{r}
ranks_test <- recompute_ranks(res, test_fns)
```

We use the `.ranks_to_sbc_results` function as an intermediary: it formats our recomputed ranks `ranks_test` object into the shape expected by the SBC plot functions.

```{r}
ranks_res <- .ranks_to_sbc_results(ranks_test)
```

Then we can call the SBC plot functions on the appropriately formatted `ranks_res` object:

```{r}
SBC::plot_rank_hist(ranks_res)
SBC::plot_ecdf_diff(ranks_res)
```

## User workflow

In conclusion, the user workflow is as follows:

1. You fit the model ONCE.

  result <- run_SBC(...)
  
2. Plot baseline parameters - use SBC directly from the package.

  SBC::plot_rank_hist(result$sbc_result)
  SBC::plot_ecdf(result$sbc_result)
  SBC::plot_ecdf_diff(result$sbc_result)
  
3. Define test functions

  test_fns <- make_test_functions(...)
  
4. Recompute ranks, change into appropriate format, and plot

  ranked <- recompute_ranks(result, test_fns)
  
  ranked_res <- .ranks_to_sbc_results(ranks_test)
  
  SBC::plot_rank_hist(ranked_res)
  SBC other plots, etc.
  
  






