---
title: "test_Geweke"
format: html
---

## Testing Geweke
### On the iid normal example

Loading dependencies:

```{r}
#| label: load-packages
library(SBC)
```

Load our package:

```{r}
#| label: load-our-code

source("R/sbc.R")
source("R/geweke.R")
```

Let's define our three sampler functions (taken directly from `tutorial_bayes_gibbs_geweke`):

```{r}
#| label: correct-three-samplers

prior_sampler <- function(prior_hyper_params) {
  
  x <- prior_hyper_params$x
  bbar0 <- prior_hyper_params$bbar0
  v0 <- prior_hyper_params$v0
  a0 <- prior_hyper_params$a0
  b0 <- prior_hyper_params$b0
  
  # simulate beta ~ N(bbar0, v0)
  beta <- rmvnorm(1, bbar0, v0)
  
  # independently simulate sigma ~ IG(a0, b0)
  sigma <- 1 / rgamma(1, shape = a0, rate = b0)

  return(c(setNames(as.vector(beta), paste0("beta", 1:p)), sigma = sigma))
}

likelihood_sampler <- function(n_obs, theta, x) {

  p <- ncol(x)
  
  beta <- theta[1:p]
  sigma <- theta[p + 1]
  
  #Mistakes: as.vector and rmvnorm taking 1 draw (not n) = bugs Claude caught
  y <- as.vector(rmvnorm(1, x%*%beta, sigma*diag(n)))
  
  return(y)
}

likelihood_sampler_wrapped <- function(n_obs, theta) {
  likelihood_sampler(n_obs = n_obs, theta = theta, prior_hyper_params$x)
}

posterior_sampler <- function(ndraws, y, prior_hyper_params) {
                            # init = c(mean(y), var(y))){ #need to adjust init
  
  x <- prior_hyper_params$x
  bbar0 <- prior_hyper_params$bbar0
  v0 <- prior_hyper_params$v0
  a0 <- prior_hyper_params$a0
  b0 <- prior_hyper_params$b0
  
  # data summaries
  n <- length(y)
  p <- ncol(x) #input type alert: need to input X as a matrix not a vector
  
  v0_inv <- solve(v0)
  t_x <- t(x)

  # preallocate storage and initialize 
  
  THETA <- matrix(0, nrow = ndraws, ncol = p+1)

  theta <- c(rep(0, p), var(y)) #init

  for(s in 1:ndraws){
    
    # draw beta given sigma and data
    
    v_n <- solve(v0_inv + t_x%*%x/theta[p+1])
    bbar_n <- v_n %*% (v0_inv%*%bbar0 + t_x%*%y/theta[p+1]) 
    #where to use bbar0 and how to initialize
    
    theta[1:p] <- rmvnorm(1, bbar_n, v_n) #check if this works as intended
    
    # draw sigma given beta and data
    
    a_n <- n/2 + a0
    b_n <- b0 + 0.5*(t(y)%*%y - 
      2*t(theta[1:p])%*%t_x%*%y + 
      t(theta[1:p])%*%t_x%*%x%*%theta[1:p])
    
    #resid <- as.vector(y - x%*%theta[1:p])
    #b_n <- b0 + as.numeric(t(resid)%*%resid)*0.5
    
    
    theta[p+1] <- 1 / (rgamma(1, shape = a_n, rate = b_n))
    
    #store at end of each iteration of the loop
    
    THETA[s, ] <- theta
    
  }
  
  colnames(THETA) <- c(paste0("beta", 1:p), "sigma")
  return(THETA)
}
```

```{r}
#| label: initialize-prior-hyper-params

p <- 3
n_obs <- 10
n_draws <- 2000

x <- matrix(rnorm(n*p, 0, 1), nrow = n, ncol = p)

prior_hyper_params <- list(x = x, 
                           a0 = 1, 
                           b0 = 1, 
                           bbar0 = rep(0, p), 
                           v0 = diag(p))
```


Get our draws:

```{r}
#| label: run-geweke-functions

#get the direct joint sampler using marginal-conditional

library(mvtnorm)

direct_draws <- geweke_mc_draws(prior_sampler = prior_sampler, 
                                likelihood_sampler = likelihood_sampler_wrapped,
                                prior_hyper_params = prior_hyper_params,
                                n_draws = n_draws, n_obs = n_obs)

#get gibbs joint sampler using successive-conditional

gibbs_draws <- geweke_sc_draws(prior_sampler = prior_sampler, 
                               likelihood_sampler = likelihood_sampler_wrapped,
                               posterior_sampler = posterior_sampler,
                               prior_hyper_params = prior_hyper_params,
                               n_draws = n_draws, n_obs = n_obs)
```

We can see our draws in both a long format (compatible with SBC) or the standard format. 
```{r}

direct_draws$draws
gibbs_draws$draws
```

Great! We have our draws. Now let's do all of our test functions to compare them quantitatively and qualitatively.

Time to define our test functions in a specific way. Like this:

```{r}
#| label: define-test-functions
#| eval: false

test_fns <- make_test_functions(
  mu_sq = function(theta, y) theta["mu"]^2,
  product = function(theta, y) theta["mu"] * theta["sigma2"],
  log_like = function(theta, y) sum(dnorm(y, 
                                          mean = theta["mu"], 
                                          sd = sqrt(theta["sigma2"]),
                                          log = TRUE))
)
```

Quantitative comparison using `tabulate_geweke_tests`:

```{r}
#| label: geweke-test-chart
#| eval: false

tabulate_geweke_tests(direct_draws = direct_draws,
                 gibbs_draws = gibbs_draws,
                 test_functions = test_fns)
```
The output is a table! Cute.

Make some QQ plots:

```{r}
#| label: geweke-test-qq-plots
#| eval: false

plot_geweke_tests(direct_draws = direct_draws,
                 gibbs_draws = gibbs_draws,
                 test_functions = test_fns)
```


For visual comparison, we made the outputs compatible with SBC plotting functions. But methodologically. Gibbs sampling isn't really built for rank computation, so it's not a meaningful result.

```{r}
#| label: plot-geweke-on-sbc-graphs
#| eval: false

SBC::plot_rank_hist(gibbs_draws$draws)
SBC::plot_ecdf(gibbs_draws$draws)
SBC::plot_ecdf_diff(gibbs_draws$draws)
```


## User workflow

In conclusion, the user workflow is as follows:

1. You fit the model ONCE.

  result <- run_SBC(...)
  
2. Plot baseline parameters - use SBC directly from the package.

  SBC::plot_rank_hist(result$sbc_result)
  SBC::plot_ecdf(result$sbc_result)
  SBC::plot_ecdf_diff(result$sbc_result)
  
3. Define test functions

  test_fns <- make_test_functions(...)
  
4. Recompute ranks, change into appropriate format, and plot

  ranked <- recompute_ranks(result, test_fns)
  
  ranked_res <- .ranks_to_sbc_results(ranks_test)
  
  SBC::plot_rank_hist(ranked_res)
  SBC other plots, etc.
  
  






