---
title: "test_distributional_shift"
format: html
---

## Testing distributional shift analysis

### On the iid normal–inverse-gamma example

Load the package modules:

```{r}
source("R/geweke.R")
source("R/testfunctions.R")
source("R/DistributionShift.R")
```

------------------------------------------------------------------------

## 1. Define the model

Same conjugate normal–inverse-gamma model used in `test_Geweke_iid_normal.qmd`.

```{r}
prior_hyper_params <- c(a0 = 1, b0 = 1, mu0 = 0, v20 = 1)

prior_sampler <- function(prior_hyper_params) {
  a0  <- prior_hyper_params[["a0"]]
  b0  <- prior_hyper_params[["b0"]]
  mu0 <- prior_hyper_params[["mu0"]]
  v20 <- prior_hyper_params[["v20"]]
  mu     <- rnorm(1, mu0, sqrt(v20))
  sigma2 <- 1 / rgamma(1, a0, rate = b0)
  c(mu = mu, sigmasq = sigma2)
}

likelihood_sampler <- function(n, theta) {
  rnorm(n, mean = theta[["mu"]], sd = sqrt(theta[["sigmasq"]]))
}

posterior_sampler <- function(ndraws, y, prior_hyper_params,
                               init = c(mean(y), var(y))) {
  n     <- length(y)
  sum_y <- sum(y)
  a0    <- prior_hyper_params[["a0"]]
  b0    <- prior_hyper_params[["b0"]]
  mu0   <- prior_hyper_params[["mu0"]]
  v20   <- prior_hyper_params[["v20"]]

  THETA <- matrix(0, nrow = ndraws, ncol = 2,
                  dimnames = list(NULL, c("mu", "sigmasq")))
  theta <- init

  for (s in seq_len(ndraws)) {
    v2n      <- 1 / (1 / v20 + n / theta[2])
    mun      <- v2n * (mu0 / v20 + sum_y / theta[2])
    theta[1] <- rnorm(1, mun, sqrt(v2n))

    an       <- a0 + n / 2
    bn       <- b0 + sum((y - theta[1])^2) / 2
    theta[2] <- 1 / rgamma(1, shape = an, rate = bn)

    THETA[s, ] <- theta
  }
  THETA
}
```

------------------------------------------------------------------------

## 2. Generate the two draw populations

```{r}
set.seed(42)

direct_draws <- geweke_mc_draws(
  prior_sampler      = prior_sampler,
  likelihood_sampler = likelihood_sampler,
  prior_hyper_params = prior_hyper_params,
  n_draws            = 2000,
  n_obs              = 50
)

gibbs_draws <- geweke_sc_draws(
  prior_sampler      = prior_sampler,
  likelihood_sampler = likelihood_sampler,
  posterior_sampler  = posterior_sampler,
  prior_hyper_params = prior_hyper_params,
  n_draws            = 2000,
  n_obs              = 50
)
```

------------------------------------------------------------------------

## 3. Run distributional shift analysis — correct sampler

For a correct sampler the two populations come from the same joint distribution p(θ, y), so the classifier accuracy should be near 0.50 and all divergences near zero.

```{r}
shift_res <- run_distributional_shift(
  direct_draws = direct_draws,
  gibbs_draws  = gibbs_draws,
  k_prototypes = 4L
)
```

Synthesis table — classifier importances joined with divergence metrics:

```{r}
shift_res$synthesis
```

Divergence metrics table (Method 3):

```{r}
tabulate_shift_divergences(shift_res)
```

------------------------------------------------------------------------

## 4. Visualise the three methods — correct sampler

Method 1: bars should all be near zero.

```{r}
plot_shift_importance(shift_res)
```

Method 2: bar heights should be similar for A and B at every prototype.

```{r}
plot_shift_prototypes(shift_res)
```

Method 3: blue and red curves should overlap closely.

```{r}
plot_shift_densities(direct_draws, gibbs_draws, shift_res)
```

------------------------------------------------------------------------

## 5. Deliberately broken sampler

Replace the Gibbs sampler with one that never updates `mu` — the variance block runs correctly but the mean is stuck at its initialisation value.

```{r}
broken_posterior <- function(ndraws, y, prior_hyper_params,
                              init = c(mean(y), var(y))) {
  n  <- length(y)
  a0 <- prior_hyper_params[["a0"]]
  b0 <- prior_hyper_params[["b0"]]

  THETA <- matrix(0, nrow = ndraws, ncol = 2,
                  dimnames = list(NULL, c("mu", "sigmasq")))
  theta <- init

  for (s in seq_len(ndraws)) {
    # BUG: mu is never updated
    an       <- a0 + n / 2
    bn       <- b0 + sum((y - theta[1])^2) / 2
    theta[2] <- 1 / rgamma(1, shape = an, rate = bn)
    THETA[s, ] <- theta
  }
  THETA
}

gibbs_broken <- geweke_sc_draws(
  prior_sampler      = prior_sampler,
  likelihood_sampler = likelihood_sampler,
  posterior_sampler  = broken_posterior,
  prior_hyper_params = prior_hyper_params,
  n_draws            = 2000,
  n_obs              = 50
)

shift_broken <- run_distributional_shift(
  direct_draws = direct_draws,
  gibbs_draws  = gibbs_broken,
  k_prototypes = 4L
)
```

`mu` should dominate the importance ranking with a large divergence; `sigmasq` should look fine:

```{r}
shift_broken$synthesis
```

```{r}
tabulate_shift_divergences(shift_broken)
```

```{r}
plot_shift_importance(shift_broken,
                      label_A = "Direct (MC)",
                      label_B = "Broken Gibbs")
```

```{r}
plot_shift_prototypes(shift_broken,
                      label_A = "Direct (MC)",
                      label_B = "Broken Gibbs")
```

```{r}
plot_shift_densities(direct_draws, gibbs_broken, shift_broken,
                     label_A = "Direct (MC)",
                     label_B = "Broken Gibbs")
```

------------------------------------------------------------------------

## User workflow summary

```         
source("R/geweke.R")
source("R/testfunctions.R")
source("R/DistributionShift.R")

# 1. Generate draw populations
direct_draws <- geweke_mc_draws(...)
gibbs_draws  <- geweke_sc_draws(...)

# 2. Confirm Geweke passes first
plot_geweke_tests(direct_draws, gibbs_draws, test_fns)
tabulate_geweke_tests(direct_draws, gibbs_draws, test_fns)

# 3. Explain any difference
shift_res <- run_distributional_shift(direct_draws, gibbs_draws)

# 4. Inspect numerically
shift_res$synthesis
tabulate_shift_divergences(shift_res)

# 5. Plot
plot_shift_importance(shift_res)
plot_shift_prototypes(shift_res)
plot_shift_densities(direct_draws, gibbs_draws, shift_res)
```
