---
title: "test_Geweke"
format: html
---

## Testing Geweke
### On the iid normal example

Loading dependencies:

```{r}
#| label: load-packages
library(SBC)
```

Load our package:

```{r}
#| label: load-our-code

source("R/sbc.R")
source("R/geweke.R")
```

Let's define our three sampler functions (taken directly from `tutorial_bayes_gibbs_geweke`):

```{r}
#| label: correct-three-samplers

prior_sampler <- function(prior_hyper_params) {
  
  a0 <- prior_hyper_params$a0
  b0 <- prior_hyper_params$b0
  mu0 <- prior_hyper_params$mu0
  v20 <- prior_hyper_params$v20
  
  # simulate mu ~ N(mu0, v20)
  mu <- rnorm(1, mu0, sqrt(v20))
  
  # independently simulate sigma2 ~ IG(a0, b0)
  sigma2 <- 1 / rgamma(1, a0, rate = b0)

  return(c(mu, sigma2))
}

likelihood_sampler <- function(n, theta) {
  
  mu <- theta[1]
  sigma2 <- theta[2]
  
  y <- rnorm(n, mu, sqrt(sigma2))
  
  return(y)
}

posterior_sampler <- function(ndraws, y, prior_hyper_params,
                              init = c(mean(y), var(y))) {

  a0 <- prior_hyper_params$a0
  b0 <- prior_hyper_params$b0
  mu0 <- prior_hyper_params$mu0
  v20 <- prior_hyper_params$v20
  n <- length(y)

  # preallocate storage
  THETA <- matrix(0, nrow = ndraws, ncol = length(init))
  
  # initialize
  theta <- init

  for(s in 1:ndraws){
    
    # draw mean given variance and data
    
    v2n = 1 / (1/v20 + n/theta[2]) 
    mun = v2n * (mu0/v20 + sum(y)/theta[2])
    
    theta[1] <- rnorm(1, mun, sqrt(v2n))
    
    # draw variance given mean and data
    
    an = a0 + n/2
    bn = b0 + sum((y - theta[1])^2) / 2
    
    theta[2] <- 1 / rgamma(1, shape = an, rate = bn)
    
    # store the latest draw in the sth row of the THETA matrix
    
    THETA[s, ] <- theta
    
  }
  
  return(THETA)
}
```

Now that we have these functions, we can use them to simulate the Bayesian joint distribution two ways:

```{r}
#| label: correct-direct-join-sampler

joint_sampler_direct <- function(ndraws, n, prior_hyper_params,
                                 init = c(0, 1)) {
  
  a0 <- prior_hyper_params$a0
  b0 <- prior_hyper_params$b0
  mu0 <- prior_hyper_params$mu0
  v20 <- prior_hyper_params$v20
  
  # preallocate storage for all of the draws
  
  direct_draws <- matrix(0, ndraws, n + 2)
  colnames(direct_draws) <- c("mu", "sigma2", paste("y", 1:n, sep = ""))
  
  for(m in 1:ndraws){
    
    # simulate exactly from the prior
    
    theta <- prior_sampler(list(a0 = a0, b0 = b0, mu0 = mu0, v20 = v20))
    
    # simulate exactly from the likelihood given the draw from the prior
    
    y <- likelihood_sampler(n, theta)
    
    direct_draws[m, ] <- c(theta, y)
    
  }
  
  # return an iid sample from the joint distribution
  
  return(direct_draws)
}

```

```{r}
#| label: correct-gibbs-joint-sampler

joint_sampler_gibbs <- function(ndraws, n, prior_hyper_params) {
  
  a0 <- prior_hyper_params$a0
  b0 <- prior_hyper_params$b0
  mu0 <- prior_hyper_params$mu0
  v20 <- prior_hyper_params$v20
  
  # preallocate storage for all of the draws
  
  gibbs_draws <- matrix(0, ndraws, n + 2)
  colnames(gibbs_draws) <- c("mu", "sigma2", paste("y", 1:n, sep = ""))
  
  # initialize
  
  theta <- c(0, 1)
  y <- runif(n)
  
  for(m in 1:ndraws){
    
    # simulate from the posterior using my sampler from above
    
    # the initialization is very important here!
    theta <- posterior_sampler(1, y, list(a0 = a0, b0 = b0, mu0 = mu0, v20 = v20), 
                               init = theta)
    theta <- c(theta)
    
    # simulate data given the latest parameter values
    
    y <- likelihood_sampler(n, theta)
    
    gibbs_draws[m, ] <- c(theta, y)
    
  }
  
  return(gibbs_draws)
}

```

Set hyperparameters:

```{r}
#| label: define-hyperparams

prior_hyper_params <- list(a0 = 1, b0 = 1, mu0 = 0, v20 = 1)
```

Get our draws:

```{r}
#| label: run-geweke-functions

#get the direct joint sampler using marginal-conditional

direct_draws <- geweke_mc_draws(prior_sampler = prior_sampler, 
                                likelihood_sampler = likelihood_sampler,
                                prior_hyper_params = prior_hyper_params,
                                n_draws = 100, n_obs = 2)
```

#get gibbs joint sampler using successive-conditional
```{r}
gibbs_draws <- geweke_sc_draws(prior_sampler = prior_sampler, 
                               likelihood_sampler = likelihood_sampler,
                               posterior_sampler = posterior_sampler,
                               prior_hyper_params = prior_hyper_params,
                               n_draws = 100, n_obs = 2)
```

We can see our draws in both a long format (compatible with SBC) or the standard format. 
```{r}

direct_draws$draws
gibbs_draws$draws
```

Great! We have our draws. Now let's do all of our test functions to compare them quantitatively and qualitatively.

```{r}
source("R/testfunctions.R")
```

Time to define our test functions in a specific way. Like this:

```{r}
#| label: define-test-functions

test_fns <- make_test_functions(
  mu_sq = function(theta, y) theta["mu"]^2,
  product = function(theta, y) theta["mu"] * theta["sigmasq"],
  log_like = function(theta, y) sum(dnorm(y, 
                                          mean = theta["mu"], 
                                          sd = sqrt(theta["sigmasq"]),
                                          log = TRUE))
)
```

Quantitative comparison using `all_geweke_tests`:

```{r}
#| label: geweke-test-chart

all_geweke_tests(direct_draws = direct_draws,
                 gibbs_draws = gibbs_draws,
                 test_fns = test_fns)
```
The output is a table! Cute.

For visual comparison, we made the outputs compatible with SBC plotting functions:

```{r}
#| label: plot-geweke-on-sbc-graphs

SBC::plot_rank_hist(...)
SBC::plot_ecdf(...)
SBC::plot_ecdf_diff(...)
```


## User workflow

In conclusion, the user workflow is as follows:

1. You fit the model ONCE.

  result <- run_SBC(...)
  
2. Plot baseline parameters - use SBC directly from the package.

  SBC::plot_rank_hist(result$sbc_result)
  SBC::plot_ecdf(result$sbc_result)
  SBC::plot_ecdf_diff(result$sbc_result)
  
3. Define test functions

  test_fns <- make_test_functions(...)
  
4. Recompute ranks, change into appropriate format, and plot

  ranked <- recompute_ranks(result, test_fns)
  
  ranked_res <- .ranks_to_sbc_results(ranks_test)
  
  SBC::plot_rank_hist(ranked_res)
  SBC other plots, etc.
  
  






