---
title: "Example linear regression"
format: html
---

We have a linear regression model:
$$
y_i = x_i ^T \beta + \epsilon_i , \hspace{1cm} \epsilon_i \sim N(0, \sigma^2)
$$ 
Our unknown parameters are $\theta = \{\beta, \sigma^2\}$. (Keep in mind $\beta$ is a vector here and it follows a multivariate normal.)
The prior is given by:

$$
\beta \sim N_p(\bar{\beta}_0, V_0)
$$

$$
\sigma^2 \sim IG(a_0, b_0)
$$ 
With hyperparameters $\bar{\beta}_0, V_0, a_0, b_0$.

As before, we use marginal conditionals to approximate the posterior via Gibbs sampling: 
$$
\{\beta | \sigma^2, y, X\} \sim N_p \left(\left(V_0 ^{-1}+\frac{X^TX}{\sigma^2}\right)^{-1} \left(V_0^{-1}\beta_0 + \frac{X^T y}{\sigma^2}\right) ,\left(V_0 ^{-1}+\frac{X^TX}{\sigma^2}\right)^{-1}\right)  \newline
$$ 
$$
\{\sigma^2 | \beta, y, X\}\sim IG(a_0+\frac{n}{2}, \frac{1}{2}[\text{SSR}(\beta)]+b_0)
$$
This allows us to approximate the posterior for this model. Now onto implementing validation methods using our package.

Load our package:

```{r}
source("~/bayescheckr/R/sbc.R")
source("~/bayescheckr/R/geweke.R")
source("~/bayescheckr/R/testfunctions.R")
```

A list of (exported) functions in the package is shown:

```{r}
ls(asNamespace("bayescheckr"))
```


Hyperparameters are packed into a single object (a list), because they have different dimensions, so this is a better approach than a named vector.

Note that the constant design matrix X (predictor space data) is also included in this list. 

```{r}
prior_hyper_params <- list(
  beta_0 = c(0,0),
  v_0 = diag(2),
  a_0 = 1,
  b_0 = 1,
  X = matrix(rnorm(5 * 2), nrow = 5, ncol = 2)
)
```

Now we code up our samplers to input to the `bayescheckr` functions. Notes for the sampler function formatting: the posterior sampler MUST take an init argument to initialize theta, and the likelihood sampler MUST only take the n_obs and theta arguments. 

The prior is:

```{r}
#| label: prior_sampler

prior_sampler <- function(prior_hyper_params){
  
  beta_0 <- prior_hyper_params$beta_0
  v_0    <- prior_hyper_params$v_0
  a_0    <- prior_hyper_params$a_0
  b_0    <- prior_hyper_params$b_0
  
  #simulate betas
  beta <- as.vector(mvtnorm::rmvnorm(1, beta_0, v_0))
  
  #simulate sigma squared
  sig2 <- 1 / rgamma(1, shape = a_0, rate = b_0)
  
  return(c(beta1 = beta[1], beta2 = beta[2], sigmasq = sig2)) 
}
```

^with minimal modifications to unpack a single `prior_hyper_params` object and to return a named vector instead of unnamed.

Now the likelihood. Note, we must make a wrapper to include X (the data matrix) since our package requires that the likelihood sampler takes in only n / n_obs and theta as inputs.

```{r}
#| label: likelihood_sampler

likelihood_sampler <- function(n_obs, theta, X){
  
  #note: We have to put X in prior_hyper_params and unpack it here.
  #This is because the contract in our package is that likelihood_sampler gets
  #only n and theta as inputs.
  
  p <- ncol(X)
  
  p <- length(theta) - 1
  
  beta <- theta[1:p]
  sig2 <- theta[p+1]
  
  y <- as.numeric(X %*% beta) + rnorm(n_obs, mean = 0, sd = sqrt(sig2))
  
  return(y)
}

likelihood_sampler_wrapped <- function(n_obs, theta) {
  likelihood_sampler(n_obs = n_obs, theta = theta, prior_hyper_params$X)
}
```

And finally our Gibbs posterior sampler:

```{r}
#| label: posterior-sampler-gibbs

posterior_sampler <- function(ndraws, y, prior_hyper_params, init = c(rep(0, p), 1))
  {
  
  # unpack everything
  X      <- prior_hyper_params$X
  beta_0 <- prior_hyper_params$beta_0
  v_0    <- prior_hyper_params$v_0
  a_0    <- prior_hyper_params$a_0
  b_0    <- prior_hyper_params$b_0
  
  #pre-allocate storage and initialize
  p <- ncol(X)
  n_obs <- length(y)
  THETA <- matrix(0, nrow = ndraws, ncol = p+1)
  colnames(THETA) <- c("beta1", "beta2", "sigmasq")
  
  #initialize values
  theta <- init
  
  v_0_inv <- solve(v_0)
  X_trans <- t(X)
  
  for(s in 1:ndraws){
    
    beta_n <- solve( v_0_inv + (X_trans%*%X)/theta[p+1] ) %*% 
      ( v_0_inv%*%beta_0 + (X_trans%*%y)/theta[p+1] )
    
    v_n <- solve( v_0_inv + (X_trans%*%X)/theta[p+1] )
    
    theta[1:p] <- mvtnorm::rmvnorm(1, beta_n, v_n)
    
    #draw from updated sigma
    ssr <- t(y - X%*%theta[1:p])%*%(y - X%*%theta[1:p])
    
    a_n <- a_0 + n_obs/2
    b_n <- ssr/2 + b_0
    
    theta[p+1] <- 1 / rgamma(1, shape = a_n, rate = b_n)
    
    THETA[s, ] <- theta
    
  }
  return(THETA)
}
```

## Running it

### Geweke checks

Let's set:

```{r}
n_obs <- 5
n_draws <- 10000
```

The function `run_geweke` creates both direct and Gibbs draws from the joint.

```{r}
gibbs_results <- run_geweke(
  prior_sampler = prior_sampler,
  likelihood_sampler = likelihood_sampler_wrapped,
  posterior_sampler = posterior_sampler,
  n_obs = n_obs,
  n_draws = n_draws,
  n_burn = 1000,
  n_thin = 5,
  prior_hyper_params = prior_hyper_params,
  parallelize = TRUE
)
```

### SBC checks

The `run_sbc` function computes rank of draws from the posterior and lets us check for uniformity:

```{r}
sbc_results <- run_sbc(
  prior_sampler = prior_sampler,
  likelihood_sampler = likelihood_sampler_wrapped,
  posterior_sampler = posterior_sampler,
  n_obs = n_obs,
  n_draws = 1000,
  n_sims = 500,
  prior_hyper_params = prior_hyper_params,
  parallelize = TRUE
)
```
Let's check the marginals:
```{r}
SBC::plot_rank_hist(sbc_results$sbc_result)
SBC::plot_ecdf(sbc_results$sbc_result)
SBC::plot_ecdf_diff(sbc_results$sbc_result)
```

As expected, all fitting within the blue bands.

## Test functions

Make some test functions:

```{r}
test_fns <- make_test_functions(
  product = function(theta, y) theta["beta1"] * theta["beta2"] * theta["sigmasq"],
  log_like = function(theta, y) {
              X    <- prior_hyper_params$X
              p    <- length(theta) - 1
              beta <- theta[1:p]
              sig2 <- theta["sigmasq"]
              sum(dnorm(y, mean = X %*% beta, sd = sqrt(sig2), log = TRUE))
},
  betas_distance = function(theta, y) abs(theta["beta1"] - theta["beta2"])
)
```

### Geweke on test functions

```{r}
tabulate_geweke_tests(direct_draws = gibbs_results$direct_draws,
                 gibbs_draws = gibbs_results$gibbs_draws,
                 test_functions = test_fns)
```

```{r}
plot_geweke_tests(direct_draws = gibbs_results$direct_draws,
                 gibbs_draws = gibbs_results$gibbs_draws,
                 test_functions = test_fns)
```
### SBC on test functions

```{r}
ranks_test <- recompute_ranks(sbc_results, test_fns)

ranks_res <- ranks_to_sbc_results(ranks_test)
```

```{r}
SBC::plot_rank_hist(ranks_res)
SBC::plot_ecdf(ranks_res)
SBC::plot_ecdf_diff(ranks_res)
```

