---
title: "Example iid normal"
format: html
---
set seed first
```{r}
set.seed(42)
```

Load our package:

```{r}
#| label: load-our-code

devtools::load_all()
```
## Problem setup

This toy example is the iid normal, with the model being `n` iid draws:
$$
y_i \sim \text{N}(\mu, \,\sigma^2)
$$
With priors:
$$
\mu \sim \text{N}(\mu_0,\,v^2_0)
\newline
\sigma^2 \sim \text{IG}(a_0,\,b_0)
$$
Let $\theta = \{\mu,\,\sigma^2\}$ represent the model's unobservables. We cannot compute the posterior distribution, $p(\theta \mid y)$, directly. So we approximate it via Gibbs sampling, ie, sampling from the full conditionals:
$$
p(\mu \mid \sigma^2,\,y)
\newline
p(\sigma^2 \mid \mu, \, y)
$$

After implementing this posterior sampler, we use the functions in our package `bayescheckr` to validate them by Geweke's and SBC's frameworks.

## Implement samplers

Add our samplers:

```{r}
prior_sampler <- function(prior_hyper_params) {
  a0  <- prior_hyper_params$a0
  b0  <- prior_hyper_params$b0
  mu0 <- prior_hyper_params$mu0
  v20 <- prior_hyper_params$v20
  list(
    mu      = rnorm(1, mu0, sqrt(v20)),
    sigmasq = 1 / rgamma(1, a0, rate = b0)
  )
}
```

```{r}
likelihood_sampler <- function(n_obs, theta, prior_hyper_params) {
  
  rnorm(n_obs, theta$mu, sqrt(theta$sigmasq))
  
}
```

```{r}
posterior_sampler <- function(n_draws, y, prior_hyper_params,
                              init = c(mean(y), var(y))) {

   n <- length(y); sum.y <- sum(y)
  a0  <- prior_hyper_params$a0
  b0  <- prior_hyper_params$b0
  mu0 <- prior_hyper_params$mu0
  v20 <- prior_hyper_params$v20

  THETA <- matrix(0, nrow = n_draws, ncol = 2)
  
  theta <- c(mean(y), var(y))

  for (s in 1:n_draws) {
    v2n <- 1 / (1/v20 + n/theta[2])
    mun <- v2n * (mu0/v20 + sum.y/theta[2])
    theta[1] <- rnorm(1, mun, sqrt(v2n))

    an <- a0 + n/2
    bn <- b0 + sum((y - theta[1])^2) / 2
    theta[2] <- 1 / rgamma(1, shape = an, rate = bn)

    THETA[s, ] <- theta
  }

  # convert matrix -> list of named theta-lists (new contract)
  lapply(seq_len(n_draws), function(s)
    list(mu = THETA[s, 1], sigmasq = THETA[s, 2]))
}
```

Set hyper-parameters and initialize other values of interest:

```{r}
#| label: define-hyperparams

prior_hyper_params <- list(a0 = 1, b0 = 1, mu0 = 0, v20 = 1)

n_draws <- 5e3
n_obs <- 10
```

## Geweke

Get posterior draws by calling `run_geweke()`:

```{r}

draws <- run_geweke(prior_sampler = prior_sampler,
                    likelihood_sampler = likelihood_sampler,
                    posterior_sampler = posterior_sampler,
                    n_obs = n_obs, n_draws = n_draws, n_thin = 2,
                    prior_hyper_params = prior_hyper_params)
```

We can see our draws in both a long format (compatible with SBC) or the standard format. 
```{r}

direct_draws <- draws$direct
gibbs_draws <- draws$gibbs


direct_draws$draws
gibbs_draws$draws
```

Great! We have our draws. Now let's do all of our test functions to compare them quantitatively and qualitatively.

Time to define our test functions in a specific way. Like this:

```{r}
test_fns <- make_test_functions(
  mu_marg = function(theta, y) theta$mu,
  sigmasq_marg = function(theta, y) theta$sigmasq,
  mu_sq    = function(theta, y) theta$mu^2,
  prod = function(theta, y) theta$mu * theta$sigmasq,
  log_like = function(theta, y) sum(dnorm(y, theta$mu, sqrt(theta$sigmasq), log = TRUE))
)
```

Quantitative comparison using `tabulate_geweke_tests`:

```{r}
#| label: geweke-test-chart

tabulate_geweke_tests(direct_draws = direct_draws,
                 gibbs_draws = gibbs_draws,
                 test_functions = test_fns)
```
The output is a table! Cute.

Make some QQ plots:

```{r}
#| label: geweke-test-qq-plots

plot_geweke_tests(direct_draws = direct_draws,
                 gibbs_draws = gibbs_draws,
                 test_functions = test_fns,
                 probs = seq(0.01, 0.99, 0.01))
```

OK so now we have all our Geweke stuff. Onto SBC:

## SBC

```{r}
res <- run_sbc(
  prior_sampler      = prior_sampler,
  likelihood_sampler = likelihood_sampler,
  posterior_sampler  = posterior_sampler,
  n_sims             = 200,        # small first; scale up once it works
  n_obs              = 50,
  n_draws            = 1000,       # ditto
  prior_hyper_params = prior_hyper_params,
  test_fns           = test_fns
)
```

```{r}
SBC::plot_rank_hist(res$sbc_result)
```
```{r}
SBC::plot_ecdf_diff(res$sbc_result)
```
