---
title: "Example Reversible Jump Markov Chain Monte Carlo - Andrieu & Doucet (1999)"
format: html
---

We have a linear regression model:
$$
y_i = x_i ^T \beta + \epsilon_i , \hspace{1cm} \epsilon_i \sim N(0, \sigma^2)
$$ 
Our unknown parameters are $\theta = \{\beta, \sigma^2\}$. (Keep in mind $\beta$ is a vector here and it follows a multivariate normal.)
The prior is given by:

$$
\beta \sim N_p(\bar{\beta}_0, V_0)
$$

$$
\sigma^2 \sim IG(a_0, b_0)
$$ 
With hyperparameters $\bar{\beta}_0, V_0, a_0, b_0$.

As before, we use marginal conditionals to approximate the posterior via Gibbs sampling: 
$$
\{\beta | \sigma^2, y, X\} \sim N_p \left(\left(V_0 ^{-1}+\frac{X^TX}{\sigma^2}\right)^{-1} \left(V_0^{-1}\beta_0 + \frac{X^T y}{\sigma^2}\right) ,\left(V_0 ^{-1}+\frac{X^TX}{\sigma^2}\right)^{-1}\right)  \newline
$$ 
$$
\{\sigma^2 | \beta, y, X\}\sim IG(a_0+\frac{n}{2}, \frac{1}{2}[\text{SSR}(\beta)]+b_0)
$$
This allows us to approximate the posterior for this model. Now onto implementing validation methods using our package.

Load our package:

```{r}
devtools::load_all()
```

Hyperparameters are packed into a single object (a list), because they have different dimensions, so this is a better approach than a named vector.

Note that the constant design matrix X (predictor space data) is also included in this list. 

```{r}
#| label: problem-setup

#set up using same settings as Lance

k_fixed <- 2
n       <- 64
i_idx   <- 0:(n - 1)

prior_hyper_params <- list(
  v0     = 2,     # Jeffreys' prior on sigma^2 (eq. 6)
  gamma0 = 2,
  delta2 = 1,      # expected SNR parameter, eq. (7)
  k      = k_fixed,
  k_max  = 3,
  n      = n,
  i_idx  = i_idx,
  lambda    = 0.3,         # P(periodogram proposal) in frequency update
  Lambda = 3,
  sigma_rw  = 1 / (5*n), # random-walk sd for frequency update
  n_mcmc_sweeps = 50       # number of full update sweeps per posterior draw
)

hyper <- prior_hyper_params
```

Now we code up our samplers to input to the `bayescheckr` functions. Notes for the sampler function formatting: the posterior sampler MUST take an init argument to initialize theta, and the likelihood sampler MUST only take the n_obs and theta arguments. 

We'll first define our inner helper functions:

```{r}
#| label: define-helper-functions

doucet_build_D <- function(omega_k, n) {
  k <- length(omega_k)

  i_idx <- 0:(n-1)

  if (k == 0) {
    return(matrix(0, nrow = n, ncol = 0))
  }

  D <- matrix(0, nrow = length(i_idx), ncol = 2*k)
  for (j in seq_len(k)) {
    D[, 2*j - 1] <- cos(omega_k[j]*i_idx)
    D[, 2*j]     <- sin(omega_k[j]*i_idx)
  }
  return(D)
}

#' @export
doucet_yPy <- function(omega_k, y, hyper) {
  k <- length(omega_k)
  if (k == 0) return(sum(y^2))
  D <- doucet_build_D(omega_k, length(y))
  DtD <- crossprod(D)
  Mk_inv <- ((hyper$delta2 + 1) / hyper$delta2) * DtD
  Dty <- crossprod(D, y)
  sol <- tryCatch(solve(Mk_inv, Dty), error = function(e) NULL)
  if (is.null(sol)) return(sum(y^2))
  sum(y^2) - as.numeric(crossprod(Dty, sol))
}

#' @export
doucet_inner_update <- function(omega, a, sigma2, y, prior_hyper_params) {
  hyper <- prior_hyper_params
  k <- length(omega)
  n_obs <- length(y)
  mh_lambda <- 0.3
  sigma_rw <- 1 / (5 * n_obs)

  # FFT proposal setup (same as before)
  fft_y <- fft(y)
  periodogram <- Mod(fft_y)^2
  Np <- n_obs
  periodogram <- periodogram[1:Np]
  freq_grid <- (0:(Np - 1)) * pi / Np
  period_probs <- periodogram / sum(periodogram)

  propose_fft_freq <- function() {
    bin <- sample.int(Np, size = 1, prob = period_probs)
    lo <- freq_grid[bin]
    hi <- lo + pi / Np
    runif(1, lo, min(hi, pi))
  }

  density_fft_freq <- function(w) {
    bin <- min(Np, max(1, floor(w / (pi / Np)) + 1))
    period_probs[bin] / (pi / Np)
  }

  # MH on frequencies (same as before)
  for (j in seq_len(k)) {
    omega_minus_j <- omega[-j]
    u <- runif(1)

    if (u < mh_lambda) {
      omega_prop <- propose_fft_freq()
      q_fwd <- density_fft_freq(omega_prop)
      q_bwd <- density_fft_freq(omega[j])
    } else {
      omega_prop <- rnorm(1, mean = omega[j], sd = sigma_rw)
      if (omega_prop <= 0 || omega_prop >= pi) next
      q_fwd <- dnorm(omega_prop, mean = omega[j], sd = sigma_rw)
      q_bwd <- dnorm(omega[j], mean = omega_prop, sd = sigma_rw)
    }

    omega_cand <- sort(c(omega_minus_j, omega_prop))
    omega_old <- sort(c(omega_minus_j, omega[j]))

    s_cand <- doucet_yPy(omega_cand, y, hyper)
    s_old <- doucet_yPy(omega_old, y, hyper)

    log_target_ratio <- ((n_obs + hyper$v0) / 2) *
      (log(hyper$gamma0 + s_old) - log(hyper$gamma0 + s_cand))
    log_q_ratio <- log(q_bwd) - log(q_fwd)
    log_alpha <- log_target_ratio + log_q_ratio

    if (log(runif(1)) < log_alpha) {
      omega <- omega_cand
    }
  }

  # Gibbs (same as before)
  if (k == 0) {
    Pky <- sum(y^2)
    shape_sigma2 <- (hyper$v0 + n_obs) / 2
    rate_sigma2 <- (hyper$gamma0 + Pky) / 2
    sigma2 <- 1 / rgamma(1, shape = shape_sigma2, rate = rate_sigma2)
    a <- numeric(0)
  } else {
    D <- doucet_build_D(omega, n_obs)
    DtD <- crossprod(D)
    Mk_inv <- ((hyper$delta2 + 1) / hyper$delta2) * DtD
    Mk <- solve(Mk_inv)
    m_k <- Mk %*% crossprod(D, y)
    P_k <- diag(n_obs) - D %*% Mk %*% t(D)
    Pky <- as.numeric(t(y) %*% P_k %*% y)

    shape_sigma2 <- (hyper$v0 + n_obs) / 2
    rate_sigma2 <- (hyper$gamma0 + Pky) / 2
    sigma2 <- 1 / rgamma(1, shape = shape_sigma2, rate = rate_sigma2)
    a <- as.vector(MASS::mvrnorm(1, mu = as.numeric(m_k),
                                 Sigma = sigma2 * Mk))
  }

  list(omega = omega, a = a, sigma2 = sigma2)
}
```

The prior is:

```{r}
#| label: prior_sampler

prior_sampler <- function(prior_hyper_params) {

  n <- prior_hyper_params$n
  k_max <- prior_hyper_params$k_max
  Lambda <- prior_hyper_params$Lambda
  delta2 <- prior_hyper_params$delta2
  v0 <- prior_hyper_params$v0
  gamma0 <- prior_hyper_params$gamma0

  # 1. sample k ~ truncated Poisson on {0, ..., k_max}
  repeat {
    k <- rpois(1, Lambda)
    if (k <= k_max) break
  }

  # 2. sample omega_k ~ iid Uniform(0, pi), k-length vector
  if (k > 0) {
    omega <- runif(k, 0, pi)
  } else {
    omega <- numeric(0)
  }

  # 3. sample sigma2 ~ IG(v0/2, gamma0/2)
  sigma2 <- 1 / rgamma(1, shape = (v0/2), rate = (gamma0/2))

  # 4. sample a_k | omega_k, sigma2, k ~ N_k(0, delta2 * sigma2 * (D^T D)^{-1})
  if (k > 0) {
    D <- doucet_build_D(omega, n = n)
    DtD_inv <- solve(t(D) %*% D)
    Sigma_a <- sigma2 * delta2 * DtD_inv #sigma2 value x Sigma matrix described
    a <- as.vector(mvtnorm::rmvnorm(1,
                                    mean = rep(0, 2*k), #mu = 0 vector
                                    sigma = Sigma_a)) #covariance matrix
  } else {
    a <- numeric(0) #empty (0) vector
  }

  # 5. pad to k_max and assemble named vector
  omega_padded <- rep(NA_real_, k_max)
  if (k > 0) omega_padded[1:k] <- omega

  a_padded <- rep(NA_real_, 2 * k_max)
  if (k > 0) a_padded[1:(2 * k)] <- a

  theta <- c(k = k,
             setNames(omega_padded, paste0("omega", 1:k_max)),
             setNames(a_padded, paste0("a", 1:(2 * k_max))),
             sigma2 = sigma2)

  return(theta)
}
```

^with minimal modifications to unpack a single `prior_hyper_params` object and to return a named vector instead of unnamed.

Now the likelihood. Note, we must make a wrapper to include X (the data matrix) since our package requires that the likelihood sampler takes in only n / n_obs and theta as inputs.

```{r}
#| label: likelihood_sampler

likelihood_sampler <- function(n_obs, theta, prior_hyper_params) {

  # infer k_max from vector length: 1(k) + k_max(omega) + 2*k_max(a) + 1(sigma2)
  k_max <- (length(theta) - 2) / 3

  k <- theta["k"]
  sigma2 <- theta["sigma2"]

  if (k > 0) {
    omega <- theta[paste0("omega", 1:k)]   # only first k slots, rest are NA
    a <- theta[paste0("a", 1:(2 * k))]  # only first 2k slots
    D <- doucet_build_D(omega, n_obs)
    mu <- as.vector(D %*% a)
  } else {
    mu <- rep(0, n_obs)
  }

  y <- rnorm(n_obs, mean = mu, sd = sqrt(sigma2))
  return(y)
}
```

And finally our posterior sampler:

```{r}
#| label: posterior-sampler

posterior_sampler <- function(n_draws,
                              y,
                              prior_hyper_params,
                              init = NULL,
                              buggy = FALSE #toggle on to replicate original error
                              ) {
  hyper <- prior_hyper_params
  n_obs <- length(y)
  k_max <- hyper$k_max
  Lambda <- hyper$Lambda

  # Initialize from prior or use provided init
  if (is.null(init)) {
    init_state <- doucet_prior_sampler(prior_hyper_params)
  } else {
    init_state <- init
  }
  k <- init_state["k"]
  omega <- sort(init_state[grepl("^omega", names(init_state))])
  a <- init_state[grepl("^a", names(init_state))]
  sigma2 <- init_state["sigma2"]

  c_prob <- 0.5
  b <- function(k) {
    if (k == k_max) return(0)
    c_prob * min(1, dpois(k + 1, Lambda) / dpois(k, Lambda))
  }
  d <- function(k) {
    if (k == 0) return(0)
    c_prob * min(1, dpois(k - 1, Lambda) / dpois(k, Lambda))
  }

  col_names <- c("k", paste0("omega_", 1:k_max),
                 paste0("a_c", 1:k_max), paste0("a_s", 1:k_max),
                 "sigma2")
  out <- matrix(NA_real_, nrow = n_draws, ncol = length(col_names),
                dimnames = list(NULL, col_names))

  for (i in seq_len(n_draws)) {
    bk <- b(k)
    dk <- d(k)
    u <- runif(1)

    if (u < bk) {
      omega_new <- runif(1, 0, pi)
      omega_prop <- sort(c(omega, omega_new))

      s_old <- doucet_yPy(omega, y, hyper)
      s_prop <- doucet_yPy(omega_prop, y, hyper)

      r_birth <- ((hyper$gamma0 + s_old) / (hyper$gamma0 + s_prop))^((n_obs + hyper$v0) / 2) *
        1 / (1 + hyper$delta2)
      
      #This is the mistake! Acceptance ratio for birth moves was too small, overly restricting 
      #the movement of the MCMC chain
      
      if (buggy) r_birth <- r_birth / (k + 1)

      if (runif(1) < min(1, r_birth)) {
        k <- k + 1
        omega <- omega_prop
        updated <- doucet_inner_update(omega, rep(0, 2*k), sigma2, y, hyper)
        a <- updated$a
        sigma2 <- updated$sigma2
      }

    } else if (u < bk + dk) {
      j_kill <- sample.int(k, 1)
      omega_prop <- omega[-j_kill]

      s_old <- doucet_yPy(omega, y, hyper)
      s_prop <- doucet_yPy(omega_prop, y, hyper)

      r_birth <- ((hyper$gamma0 + s_prop) / (hyper$gamma0 + s_old))^((n_obs + hyper$v0) / 2) *
        1 / (1 + hyper$delta2)

      if (runif(1) < min(1, 1 / r_birth)) {
        k <- k - 1
        omega <- omega_prop
        a <- if (k > 0) a[-c(2*j_kill - 1, 2*j_kill)] else numeric(0)
        updated <- doucet_inner_update(omega, a, sigma2, y, hyper)
        a <- updated$a
        sigma2 <- updated$sigma2
      }

    } else {
      updated <- doucet_inner_update(omega, a, sigma2, y, hyper)
      omega <- updated$omega
      a <- updated$a
      sigma2 <- updated$sigma2
    }

    omega_pad <- rep(NA_real_, k_max)
    if (k > 0) omega_pad[1:k] <- omega

    a_pad <- rep(NA_real_, 2 * k_max)
    if (k > 0) a_pad[1:(2*k)] <- a

    out[i, ] <- c(k, omega_pad, a_pad, sigma2)
  }

  return(out)
}
```

### Test functions

```{r}
#| label: test-functions

#define log-likelihood function

log_lik <- function(theta, y) {

  k <- as.integer(theta[["k"]])
  sigma2 <- theta[["sigma2"]]
  n <- length(y)

  # construct mean vector
  if (k > 0) {
    omega <- as.numeric(theta[paste0("omega", 1:k)])
    a <- as.numeric(theta[paste0("a", 1:(2 * k))])

    D <- doucet_build_D(omega, n)
    mu <- as.vector(D %*% a)
  } else {
    mu <- rep(0, n)
  }

  sum(dnorm(y,
            mean = mu,
            sd = sqrt(sigma2),
            log = TRUE))
}

a_prod = function(theta, y) {
    k <- theta[["k"]]
    if (k == 0) return(0)

    a <- theta[paste0("a", 1:(2 * k))]
    prod(a, na.rm = TRUE)
}

omega_prod = function(theta, y) {
    k <- theta[["k"]]
    if (k == 0) return(0)
    else if (is.na(k)) return(NA)
    prod(theta[paste0("omega", 1:k)], na.rm = TRUE)
  }

test_fns <- make_test_functions(k_marg = function(theta, y) theta["k"], #marginal of k
                                omega_prod = omega_prod,
                                log_likelihood = log_lik, #log likelihood defined above
                                a_prod = a_prod)
```


## Running it

## Geweke checks

Let's set:

```{r}
n_obs <- 64
n_draws <- 500
```

The function `run_geweke` creates both direct and Gibbs draws from the joint.

```{r}
#| label: run-geweke
#| warning: false
doucet_results <- bayescheckr::run_geweke(
  prior_sampler = prior_sampler,
  likelihood_sampler = likelihood_sampler,
  posterior_sampler = posterior_sampler,
  n_obs = n_obs,
  n_draws = 1e3,
  n_burn = 1000,
  n_thin = 5,
  prior_hyper_params = prior_hyper_params,
  parallelize = TRUE
)

buggy_doucet_results <- bayescheckr::run_geweke(
  prior_sampler = 
)

doucet_results$gibbs$wide
```

### Geweke on test functions

```{r}
tabulate_geweke_tests(draws = doucet_results,
                      tests_applied = FALSE,
                      test_functions = test_fns)

# Alternatively, use the tests_applied parameter if you ever need to separate
# the test function application and analysis steps:
# 
# geweke_test_matrix <- apply_test_functions(direct = doucet_results$direct,
#                                            gibbs = doucet_results$gibbs,
#                                            tests_functions = test_fns)
# 
# tabulate_geweke_tests(draws = geweke_test_matrix, tests_applied = TRUE)
```

```{r}
plot_geweke_tests(direct_draws = doucet_results$direct,
                 gibbs_draws = doucet_results$gibbs,
                 test_functions = test_fns)
```

## SBC checks

The `run_sbc` function computes rank of draws from the posterior and lets us check for uniformity:

```{r}
#| eval: false
set.seed(123)

sbc_results <- run_sbc(
  prior_sampler = prior_sampler,
  likelihood_sampler = likelihood_sampler,
  posterior_sampler = posterior_sampler,
  n_obs = n_obs,
  n_draws = 2000,
  n_sims = 500,
  test_fns = test_fns,
  prior_hyper_params = prior_hyper_params,
  parallelize = TRUE
)
```
Let's check the marginals:
```{r}
SBC::plot_rank_hist(sbc_results$sbc_result)
SBC::plot_ecdf(sbc_results$sbc_result)
SBC::plot_ecdf_diff(sbc_results$sbc_result)
```

As expected, all fitting within the blue bands.