---
title: "Example linear regression"
format: html
---

Load our package:

```{r}
source("R/sbc.r")
```

Taken from `SBC on linear regression.qmd`, in `rzmautner`'s sandbox, hyperparameters were modified into a single object (a list) to match the generality of the package:

```{r}
prior_hyper_params <- list(
  beta_0 = c(0,0),
  v_0 = diag(2),
  a_0 = 1,
  b_0 = 1,
  X = matrix(rnorm(5 * 2), nrow = 5, ncol = 2)
)
```

So our prior is:

```{r}
#| label: prior_sampler

prior_sampler <- function(prior_hyper_params){
  
  beta_0 <- prior_hyper_params$beta_0
  v_0    <- prior_hyper_params$v_0
  a_0    <- prior_hyper_params$a_0
  b_0    <- prior_hyper_params$b_0
  
  #simulate betas
  beta <- as.vector(mvtnorm::rmvnorm(1, beta_0, v_0))
  
  #simulate sigma squared
  sig2 <- 1 / rgamma(1, shape = a_0, rate = b_0)
  
  return(c(beta1 = beta[1], beta2 = beta[2], sigmasq = sig2)) 
}
```

^with minimal modifications to unpack a single `prior_hyper_params` object and to return a named vector instead of unnamed.

Now the likelihood:

```{r}
#| label: likelihood_sampler

likelihood_sampler <- function(n, theta){
  
  #note: We have to put X in prior_hyper_params and unpack it here.
  #This is because the contract in our package is that likelihood_sampler gets
  #only n and theta as inputs.
  
  X <- prior_hyper_params$X
  
  p <- length(theta) - 1
  
  beta <- theta[1:p]
  sig2 <- theta[p+1]
  
  y <- X %*% beta + rnorm(n, mean = 0, sd = sqrt(sig2))
  
  return(y)
}
```

And finally our Gibbs posterior sampler:

```{r}
#| label: posterior-sampler-gibbs

posterior_sampler <- function(ndraws, y, prior_hyper_params){
  
  # unpack everything
  X      <- prior_hyper_params$X
  beta_0 <- prior_hyper_params$beta_0
  v_0    <- prior_hyper_params$v_0
  a_0    <- prior_hyper_params$a_0
  b_0    <- prior_hyper_params$b_0
  
  #pre-allocate storage and initialize
  p <- ncol(X)
  n_obs <- length(y)
  THETA <- matrix(0, nrow = ndraws, ncol = p+1)
  colnames(THETA) <- c("beta1", "beta2", "sigmasq")
  
  #initialize values
  theta <- c(rep(0, p), 1)
  
  v_0_inv <- solve(v_0)
  X_trans <- t(X)
  
  for(s in 1:ndraws){
    
    beta_n <- solve( v_0_inv + (X_trans%*%X)/theta[p+1] ) %*% 
      ( v_0_inv%*%beta_0 + (X_trans%*%y)/theta[p+1] )
    
    v_n <- solve( v_0_inv + (X_trans%*%X)/theta[p+1] )
    
    theta[1:p] <- mvtnorm::rmvnorm(1, beta_n, v_n)
    
    #draw from updated sigma
    ssr <- t(y - X%*%theta[1:p])%*%(y - X%*%theta[1:p])
    
    a_n <- a_0 + n_obs/2
    b_n <- ssr/2 + b_0
    
    theta[p+1] <- 1 / rgamma(1, shape = a_n, rate = b_n)
    
    THETA[s, ] <- theta
    
  }
  return(THETA)
}
```
