---
title: "Example assembly error (Primiceri)"
format: html
---

## Setup

We are looking at a simple iid normal model, but with a key prior difference that changes everything.

The model is:

$$
y_i\mid \theta,\,\sigma^2 \overset{\text{iid}}{\sim}\text{N}(\theta,\,\sigma^2)
$$
With priors:

$$
\begin{aligned}
\theta&\sim\text{N}(\bar{\theta}_0,\,v_0)\\
\log\sigma^2&\sim\text{N}(a_0,\,b_0).
\end{aligned}
$$
Regular Gibbs sampling won't work here because $p(\sigma ^2 \mid \theta,\,y_i)$ is now intractable. What do we do?

Kim, Shepherd and Chib posit a new variable $s$ that makes this possible.

First, we define $u = \log(\sigma ^2)$, so that $u \sim \text{N}(a_0, \,b_0)$.

Standardize the likelihood:

$$
y_i \sim \text{N}(\theta,\,\sigma^2) \implies\frac{y_i-\theta}{\sigma}\sim\text{N}(0, 1).
$$
Let $z_i = \frac{y_i - \theta}{\sigma}$. Then

$$
z_i ^2 \sim \chi ^2(1)
\\ \implies \log z_i ^2 \sim \log \chi ^2 (1)
\\ \implies \log(y_i - \theta)^2-u \sim \log \chi^2(1)
$$
Given the log-chi-squared distribution, KSC's idea is to approximate this using a **mixture of normals**.

The idea is to find $\pi_j, \mu_j, \tau_j ^2$ such that

$$
p(y_i \mid \theta,\,u) \approx \sum_{j=1}^J \pi_j \,\cdot\text{N}(u\,;\mu_j,\,\tau_j ^2)
$$
Then, let $s_i \in \{1,\cdots,J\}$ indicate **which component is active for each observation $i$**. I.e., $p(s_i = j) = \pi_j$. Then:

$$
p(u \mid \theta, \,y,\,s_i = j) = \text{N}(u\,;\mu_j,\,\tau_j ^2) \rightarrow \text{Gaussian (fully tractable).}
$$
Then the Gibbs sampler to target the posterior $p(\theta,\,u,\,s \mid y)$ takes 2 blocks, $u$ (or $\sigma^2$) and $(\theta,\,s)$:

1. $u \mid \theta, \, s,\,y$
2. $\theta,\,s\mid \sigma^2, \,y$, targeted via:
   a. $\theta \mid \sigma^2, y$
   b. $s \mid \sigma^2,\,\theta,\,y$
   
And as we know, Primiceri's error was using 3 blocks, something like:

1. $u \mid \theta, s, y$
2. $s \mid \theta, u, y$
3. $\theta \mid u, y$

where step 3 fails to condition on $s$, violating the Gibbs requirement that each block conditions on all others.

Here are the marginal conditionals, i.e., what we need to sample from to target the posterior:

### Marginal conditionals

Using Bayes' rule, we obtain that:

$$
p(s_i \mid \theta,\,u,\,y) \propto \pi_j \cdot \exp\left(-\frac{(u - \mu_j)^2}{2\tau_j ^2}\right)
$$
The marginal on $u$ is obtained by normal-normal conjugacy:

$$
p(u \mid \theta,\,s,\,y) = \text{N}(\bar{a_n}, \bar{b_n}), \text{where } \newline
\bar{b}_n = \left(\frac{1}{b_0} + \sum_{i=1}^n \frac{1}{\tau_{s_i}^2}\right)^{-1}, \newline
\bar{a}_n = \bar{b}_n\left(\frac{a_0}{b_0} + \sum_{i=1}^n \frac{\mu_{s_i}}{\tau_{s_i}^2}\right)
$$
And finally, the marginal on $\theta$ doesn't even depend on $s$ (since $\theta$ was not involved in the KSC trick):

$$
p(\theta \mid u,\,y) = \text{N}(\bar{\theta_n}, \,v_n), \text{where }\newline
v_n = \left(\frac{1}{v_0} + \frac{n}{e^u}\right)^{-1}, \newline
\bar{\theta}_n = v_n\left(\frac{\bar{\theta}_0}{v_0} + \frac{\sum y_i}{e^u}\right).
$$
This gives us a full picture to implement and test the posterior through our frameworks.

## Implementation

We have an observable $y_i$ and unobservavbles $\theta,\,\sigma^2,\,s$. To put a prior on $s$ we use the values in Table 4 of KSC (1995):

```{r}
#| label: prior hyper params

theta_bar_0 <- 0
v_0 <- 2
a_0 <- 1
b_0 <- 1
n <- 50

#Table 4 constants:
ksc_pi <- c(0.00730, 0.10556, 0.00002, 0.04395, 0.34001, 0.24566, 0.25750)
ksc_mu  <- c(-10.12999, -3.97281, -8.56686, 2.77786, 0.61942, 1.79518, -1.08819)
ksc_tau <- c(5.79596, 2.61369, 5.17950, 0.16735, 0.64009, 0.34023, 1.26261)

prior_hyper_params_primiceri <- c(theta_bar_0 = 0, v_0 = 2, a_0 = 1, b_0 = 1, n = 10,
                        ksc_pi = ksc_pi, ksc_mu, ksc_tau)

```


```{r}
prior_sampler_primiceri <- function(prior_hyper_params){

  theta_bar_0 <- as.numeric(prior_hyper_params["theta_bar_0"])
  v_0         <- as.numeric(prior_hyper_params["v_0"])
  a_0         <- as.numeric(prior_hyper_params["a_0"])
  b_0         <- as.numeric(prior_hyper_params["b_0"])
  n           <- as.numeric(prior_hyper_params["n"])
  
  theta  <- rnorm(1, mean = theta_bar_0, sd = sqrt(v_0))
  u      <- rnorm(1, mean = a_0, sd = sqrt(b_0))
  sigmasq <- exp(u)
  s      <- sample(1:7, size = n, replace = TRUE, prob = ksc_pi)
  
  ret <- c(list(theta = theta, sigmasq = sigmasq), setNames(as.list(s), paste0("s_", seq_len(n))))
  ret
}
```

```{r}
likelihood_sampler_primiceri <- function(n_obs, theta, prior_hyper_params){
  rnorm(n_obs, mean = theta$theta, sd = sqrt(theta$sigmasq ))
}
```

And now the tricky part. The structure goes like:
1. initialize theta, u, s.
For i in 1...n_draws: 
  1. draw u | theta, s, y (normal conjugacy update using KSC)
  2. draw theta | u, y (normal conjugacy using true likelihood, no KSC)
  3. draw s_i | u, theta, y for each i (categorical draw over 7 components)
  4. i <- i+1

```{r}
correct_posterior_sampler <- function(n_draws, y, prior_hyper_params,
                                      init = NULL){
  
  # extract hyperparameters
  theta_bar_0 <- prior_hyper_params["theta_bar_0"]
  v_0         <- prior_hyper_params["v_0"]
  a_0         <- prior_hyper_params["a_0"]
  b_0         <- prior_hyper_params["b_0"]
  n           <- length(y)
  
  # preallocate storage
  THETA <- matrix(0, nrow = n_draws, ncol = 2 + n)
  colnames(THETA) <- c("theta", "sigmasq", paste0("s_", 1:n))
  
  # initialize from init if provided, otherwise use data-based defaults
  if (!is.null(init)) {
    theta_cur <- as.numeric(init["theta"])
    u_cur     <- log(as.numeric(init["sigmasq"]))
    s_names   <- grep("^s_", names(init), value = TRUE)
    s_cur     <- if (length(s_names) == n) as.numeric(init[s_names]) else rep(5, n)
  } else {
    theta_cur <- mean(y)
    u_cur     <- log(var(y))
    s_cur     <- rep(5, n)
  }
  
  for (i in 1:n_draws) {
    
    # step 1: draw u | theta, s, y
    z     <- log((y - theta_cur)^2 + 1e-20)
    z_tilde <- z - (ksc_mu[s_cur] - 1.2704)   # mean-corrected observations
    b_n   <- 1 / (1/b_0 + sum(1/ksc_tau[s_cur]))
    a_n   <- b_n * (a_0/b_0 + sum(z_tilde / ksc_tau[s_cur]))
    u_prop <- rnorm(1, mean = a_n, sd = sqrt(b_n))
    
    # true log-likelihood at proposed and current u
    loglik_true <- function(u) {
    sum(dnorm(y, mean = theta_cur, sd = sqrt(exp(u)), log = TRUE))
    }

    # mixture approximation log-likelihood at proposed and current u
    loglik_mix <- function(u) {
      # for each observation i, compute log(sum_j pi_j * N(z_i; u + m_j - 1.2704, v_j))
      log_mix_i <- sapply(1:n, function(ii) {
        log(sum(ksc_pi * dnorm(z[ii], mean = u + ksc_mu - 1.2704, sd = sqrt(ksc_tau))))
      })
    sum(log_mix_i)
    }

    # log acceptance ratio
    log_alpha <- (loglik_true(u_prop) - loglik_true(u_cur)) -
                 (loglik_mix(u_prop)  - loglik_mix(u_cur))

    # MH step
    if (log(runif(1)) < log_alpha) {
      u_cur <- u_prop
    }
    
    # step 2: draw theta | u, y
    sigmasq_cur <- exp(u_cur)
    v_n         <- 1 / (1/v_0 + n/sigmasq_cur)
    theta_n     <- v_n * (theta_bar_0/v_0 + sum(y)/sigmasq_cur)
    theta_cur   <- rnorm(1, mean = theta_n, sd = sqrt(v_n))
    
    # step 3: draw s | u, y
    z <- log((y - theta_cur)^2 + 1e-20)  # recompute with updated theta_cur
    s_cur <- sapply(1:n, function(ii) {
    log_w <- log(ksc_pi) + dnorm(z[ii],
                                  mean = u_cur + ksc_mu - 1.2704,
                                  sd   = sqrt(ksc_tau),
                                  log  = TRUE)
    log_w <- log_w - max(log_w)
    w <- exp(log_w) / sum(exp(log_w))
    sample(1:7, size = 1, prob = w)
    })
    
    # store
    THETA[i, "theta"]            <- theta_cur
    THETA[i, "sigmasq"]          <- exp(u_cur)
    THETA[i, paste0("s_", 1:n)]  <- s_cur
  }
  
  return(THETA)
}
```

## Testing

```{r}
devtools::install_github("Bayescheck/bayescheckr")

library(bayescheckr)
```
```{r}
test_fns <- make_test_functions(
  theta = function(theta, y) theta$theta,
  
  sigmasq = function(theta, y) exp(theta$sigmasq),
  
  product = function(theta, y) {
    theta$theta * log(theta$sigmasq)
  },
  
  log_likelihood = function(theta, y) {
    sum(dnorm(y, mean = theta$theta, sd = sqrt(exp(theta$sigmasq)), log = TRUE))
  },
  
  sum_s = function(theta, y) sum(unlist(theta[grep("^s_", names(theta))])),
  
  mean_s = function(theta, y) mean(unlist(theta[grep("^s_", names(theta))]))
)
```

### Primiceri's erroneous sampler

Primiceri's erroneous posterior sampler:

```{r}
set.seed(67)

wrong_assembly_posterior_sampler <- function(n_draws, y, prior_hyper_params,
                                    init = NULL){
  
  # extract hyperparameters
  theta_bar_0 <- prior_hyper_params["theta_bar_0"]
  v_0         <- prior_hyper_params["v_0"]
  a_0         <- prior_hyper_params["a_0"]
  b_0         <- prior_hyper_params["b_0"]
  n           <- length(y)
  
  # preallocate storage
  THETA <- matrix(0, nrow = n_draws, ncol = 2 + n)
  colnames(THETA) <- c("theta", "sigmasq", paste0("s_", 1:n))
  
  # initialize from init if provided, otherwise use data-based defaults
  if (!is.null(init)) {
    theta_cur <- as.numeric(init["theta"])
    u_cur     <- log(as.numeric(init["sigmasq"]))
    s_names   <- grep("^s_", names(init), value = TRUE)
    s_cur     <- if (length(s_names) == n) as.numeric(init[s_names]) else rep(5, n)
  } else {
    theta_cur <- mean(y)
    u_cur     <- log(var(y))
    s_cur     <- rep(5, n)
  }
  
  for (i in 1:n_draws) {
    
    # step 1: draw u | theta, s, y
    z     <- log((y - theta_cur)^2 + 1e-20)
    z_tilde <- z - (ksc_mu[s_cur] - 1.2704)   # mean-corrected observations
    b_n   <- 1 / (1/b_0 + sum(1/ksc_tau[s_cur]))
    a_n   <- b_n * (a_0/b_0 + sum(z_tilde / ksc_tau[s_cur]))
    u_prop <- rnorm(1, mean = a_n, sd = sqrt(b_n))
    
    # true log-likelihood at proposed and current u
    loglik_true <- function(u) {
    sum(dnorm(y, mean = theta_cur, sd = sqrt(exp(u)), log = TRUE))
    }

    # mixture approximation log-likelihood at proposed and current u
    loglik_mix <- function(u) {
      # for each observation i, compute log(sum_j pi_j * N(z_i; u + m_j - 1.2704, v_j))
      log_mix_i <- sapply(1:n, function(ii) {
        log(sum(ksc_pi * dnorm(z[ii], mean = u + ksc_mu - 1.2704, sd = sqrt(ksc_tau))))
      })
    sum(log_mix_i)
    }

    # log acceptance ratio
    log_alpha <- (loglik_true(u_prop) - loglik_true(u_cur)) -
                 (loglik_mix(u_prop)  - loglik_mix(u_cur))

    # MH step
    if (log(runif(1)) < log_alpha) {
      u_cur <- u_prop
    }
    
    # step 2: draw s | u, theta, y
    z <- log((y - theta_cur)^2 + 1e-20)  # recompute with OUTDATED theta_cur
    s_cur <- sapply(1:n, function(ii) {
    log_w <- log(ksc_pi) + dnorm(z[ii],
                                  mean = u_cur + ksc_mu - 1.2704,
                                  sd   = sqrt(ksc_tau),
                                  log  = TRUE)
    log_w <- log_w - max(log_w)
    w <- exp(log_w) / sum(exp(log_w))
    sample(1:7, size = 1, prob = w)
    })
    
    # step 3: draw theta | u, y
    # not conditioned on s -> mistake
    sigmasq_cur <- exp(u_cur)
    v_n         <- 1 / (1/v_0 + n/sigmasq_cur)
    theta_n     <- v_n * (theta_bar_0/v_0 + sum(y)/sigmasq_cur)
    theta_cur   <- rnorm(1, mean = theta_n, sd = sqrt(v_n))
    
    # store
    THETA[i, "theta"]   <- theta_cur
    THETA[i, "sigmasq"] <- exp(u_cur)
    THETA[i, paste0("s_", 1:n)]  <- s_cur
  }
  
  return(THETA)
}
```

One posterior to rule them all:

```{r}
#| label: hybrid-posterior-primiceri

posterior_sampler <- function(n_draws,
                              y, 
                              prior_hyper_params,
                              init = NULL,
                              buggy = FALSE){
  
  # extract hyperparameters
  theta_bar_0 <- prior_hyper_params["theta_bar_0"]
  v_0         <- prior_hyper_params["v_0"]
  a_0         <- prior_hyper_params["a_0"]
  b_0         <- prior_hyper_params["b_0"]
  n           <- length(y)
  
  # preallocate storage
  THETA <- matrix(0, nrow = n_draws, ncol = 2 + n)
  colnames(THETA) <- c("theta", "sigmasq", paste0("s_", 1:n))
  
  # initialize from init if provided, otherwise use data-based defaults
  if (!is.null(init)) {
    theta_cur <- as.numeric(init["theta"])
    u_cur     <- log(as.numeric(init["sigmasq"]))
    s_names   <- grep("^s_", names(init), value = TRUE)
    s_cur     <- if (length(s_names) == n) as.numeric(init[s_names]) else rep(5, n)
  } else {
    theta_cur <- mean(y)
    u_cur     <- log(var(y))
    s_cur     <- rep(5, n)
  }
  
  for (i in 1:n_draws) {
    
    # step 1: draw u | theta, s, y
    z     <- log((y - theta_cur)^2 + 1e-20)
    z_tilde <- z - (ksc_mu[s_cur] - 1.2704)   # mean-corrected observations
    b_n   <- 1 / (1/b_0 + sum(1/ksc_tau[s_cur]))
    a_n   <- b_n * (a_0/b_0 + sum(z_tilde / ksc_tau[s_cur]))
    u_prop <- rnorm(1, mean = a_n, sd = sqrt(b_n))
    
    # true log-likelihood at proposed and current u
    loglik_true <- function(u) {
    sum(dnorm(y, mean = theta_cur, sd = sqrt(exp(u)), log = TRUE))
    }

    # mixture approximation log-likelihood at proposed and current u
    loglik_mix <- function(u) {
      # for each observation i, compute log(sum_j pi_j * N(z_i; u + m_j - 1.2704, v_j))
      log_mix_i <- sapply(1:n, function(ii) {
        log(sum(ksc_pi * dnorm(z[ii], mean = u + ksc_mu - 1.2704, sd = sqrt(ksc_tau))))
      })
    sum(log_mix_i)
    }

    # log acceptance ratio
    log_alpha <- (loglik_true(u_prop) - loglik_true(u_cur)) -
                 (loglik_mix(u_prop)  - loglik_mix(u_cur))

    # MH step
    if (log(runif(1)) < log_alpha) {
      u_cur <- u_prop
    }
    
    if (buggy) {
      # step 2: draw s | u, theta, y
    z <- log((y - theta_cur)^2 + 1e-20)  # recompute with OUTDATED theta_cur
    s_cur <- sapply(1:n, function(ii) {
    log_w <- log(ksc_pi) + dnorm(z[ii],
                                  mean = u_cur + ksc_mu - 1.2704,
                                  sd   = sqrt(ksc_tau),
                                  log  = TRUE)
    log_w <- log_w - max(log_w)
    w <- exp(log_w) / sum(exp(log_w))
    sample(1:7, size = 1, prob = w)
    })
    
    # step 3: draw theta | u, y
    # not conditioned on s -> mistake
    sigmasq_cur <- exp(u_cur)
    v_n         <- 1 / (1/v_0 + n/sigmasq_cur)
    theta_n     <- v_n * (theta_bar_0/v_0 + sum(y)/sigmasq_cur)
    theta_cur   <- rnorm(1, mean = theta_n, sd = sqrt(v_n))
    }
    
    else {
      # step 2: draw theta | u, y
      sigmasq_cur <- exp(u_cur)
      v_n         <- 1 / (1/v_0 + n/sigmasq_cur)
      theta_n     <- v_n * (theta_bar_0/v_0 + sum(y)/sigmasq_cur)
      theta_cur   <- rnorm(1, mean = theta_n, sd = sqrt(v_n))
      
      # step 3: draw s | u, y
      z <- log((y - theta_cur)^2 + 1e-20)  # recompute with updated theta_cur
      s_cur <- sapply(1:n, function(ii) {
      log_w <- log(ksc_pi) + dnorm(z[ii],
                                    mean = u_cur + ksc_mu - 1.2704,
                                    sd   = sqrt(ksc_tau),
                                    log  = TRUE)
      log_w <- log_w - max(log_w)
      w <- exp(log_w) / sum(exp(log_w))
      sample(1:7, size = 1, prob = w)
    })
    }
    
    # store
    THETA[i, "theta"]            <- theta_cur
    THETA[i, "sigmasq"]          <- exp(u_cur)
    THETA[i, paste0("s_", 1:n)]  <- s_cur
  }
  
  return(THETA)
}
```

# Geweke checks

## On baseline correct sampler

```{r}
library(dplyr)

primiceri_geweke_correct <- run_geweke(
  prior_sampler = prior_sampler_primiceri,
  likelihood_sampler = likelihood_sampler_primiceri,
  posterior_sampler = correct_posterior_sampler,
  prior_hyper_params = prior_hyper_params_primiceri,
  n_obs = 10,
  n_draws = 5000,
  n_thin = 10,
  n_burn = 1000,
  parallelize = TRUE
)
```
```{r}
plot_geweke_tests(primiceri_geweke_correct, test_functions = test_fns,
                  probs = seq(0.15, 0.85, by = 0.05))
```
```{r}
correct_direct_draws <- primiceri_geweke_correct$direct
correct_gibbs_draws <- primiceri_geweke_correct$gibbs

c2st_primiceri_correct <- run_distributional_shift(correct_direct_draws,
                                                   correct_gibbs_draws,
                                                   classifier = default_xgb_classifier())

c2st_primiceri_correct
```

### Visualization

```{r}
#| label: fit-viz-primiceri

viz_primiceri_correct <- fit_c2st_for_viz(
  correct_direct_draws$theta,
  correct_gibbs_draws$theta,
  classifier = default_xgb_classifier(nrounds = 200),
  seed = 1
)

X_mat_primiceri_correct <- as.matrix(viz_primiceri_correct$Xtest)
sum(duplicated(X_mat_primiceri_correct))
```
(The code chunk above is simply checking that there are no duplicates in the test matrix X - if there were, it would cause problems for the t-SNE method for visualization.)

```{r}
#| label: run-tsne-primiceri

set.seed(1)
library(Rtsne)
tsne_out_primiceri_correct <- Rtsne(X_mat_primiceri_correct, dims = 2, perplexity = 30,
                             verbose = FALSE, check_duplicates = FALSE)

tsne_df_primiceri <- data.frame(
  tsne1 = tsne_out_primiceri_correct$Y[, 1],
  tsne2 = tsne_out_primiceri_correct$Y[, 2],
  source = factor(viz_primiceri_correct$ytest, levels = c(0, 1), labels = c("direct (A)", "gibbs (B)"))
)
```

```{r}
#| label: plot-tsne-primiceri

library(ggplot2)

ggplot(tsne_df_primiceri, aes(x = tsne1, y = tsne2, color = source)) +
  geom_point(alpha = 0.4, size = 0.8) +
  labs(title = "t-SNE embedding of theta-space -- Primiceri corrected (baseline)") +
  theme_minimal()
```

## On wrong assembly sampler

```{r}
library(dplyr)

primiceri_geweke_wrong <- run_geweke(
  prior_sampler = prior_sampler_primiceri,
  likelihood_sampler = likelihood_sampler_primiceri,
  posterior_sampler = wrong_assembly_posterior_sampler,
  prior_hyper_params = prior_hyper_params_primiceri,
  n_obs = 10,
  n_draws = 5000,
  parallelize = TRUE
)
```
```{r}
plot_geweke_tests(primiceri_geweke_wrong, test_functions = test_fns)
```
```{r}
wrong_direct_draws <- primiceri_geweke_wrong$direct
wrong_gibbs_draws <- primiceri_geweke_wrong$gibbs

c2st_primiceri_wrong <- run_distributional_shift(wrong_direct_draws,
                                                   wrong_gibbs_draws,
                                                   classifier = default_xgb_classifier(nrounds = 100))

c2st_primiceri_wrong
```

### Visualization

```{r}
#| label: fit-viz-primiceri

viz_primiceri_wrong <- fit_c2st_for_viz(
  wrong_direct_draws$theta,
  wrong_gibbs_draws$theta,
  classifier = default_xgb_classifier(nrounds = 200),
  seed = 1
)

X_mat_primiceri_wrong <- as.matrix(viz_primiceri_wrong$Xtest)
sum(duplicated(X_mat_primiceri_wrong))
```
```{r}
#| label: run-tsne-primiceri

set.seed(1)
library(Rtsne)
tsne_out_primiceri_wrong <- Rtsne(X_mat_primiceri_wrong, dims = 2, perplexity = 30,
                             verbose = FALSE, check_duplicates = FALSE)

tsne_df_primiceri_wrong <- data.frame(
  tsne1 = tsne_out_primiceri_wrong$Y[, 1],
  tsne2 = tsne_out_primiceri_wrong$Y[, 2],
  source = factor(viz_primiceri_wrong$ytest, levels = c(0, 1), labels = c("direct (A)", "gibbs (B)"))
)
```

```{r}
#| label: plot-tsne-primiceri

ggplot(tsne_df_primiceri_wrong, aes(x = tsne1, y = tsne2, color = source)) +
  geom_point(alpha = 0.4, size = 0.8) +
  labs(title = "t-SNE embedding of theta-space -- Primiceri original error") +
  theme_minimal()
```
A comment: Geweke checks are not the best for this specific problem. We have tried increasing n_draws, n_burnin and n_thin dramatically and it did not significantly improve the plots, the Geweke statistics or the C2ST statistics.
Fortunately, our package offers multiple independent methods of validation. Below is the SBC validation implemented on this problem.

# SBC checks

## On baseline correct sampler

Now we use SBC, a different, independent validation method:

```{r}
set.seed(234)

primiceri_sbc_correct <- run_sbc(
  prior_sampler = prior_sampler_primiceri,
  likelihood_sampler = likelihood_sampler_primiceri,
  posterior_sampler = correct_posterior_sampler,
  n_sims = 500,
  n_obs = 10,
  n_draws = 2000,
  prior_hyper_params = prior_hyper_params_primiceri,
  globals = c("correct_posterior_sampler", "ksc_tau", "ksc_mu", "ksc_pi"),
  test_fns = test_fns,
  parallelize = TRUE
)
```
```{r}
SBC::plot_rank_hist(primiceri_sbc_correct$sbc_result)
SBC::plot_ecdf_diff(primiceri_sbc_correct$sbc_result)
```

## On the wrong assembly error

```{r}
primiceri_sbc_wrong <- run_sbc(
  prior_sampler = prior_sampler_primiceri,
  likelihood_sampler = likelihood_sampler_primiceri,
  posterior_sampler = wrong_assembly_posterior_sampler,
  n_sims = 500,
  n_obs = 10,
  n_draws = 2000,
  prior_hyper_params = prior_hyper_params_primiceri,
  globals = c("wrong_assembly_posterior_sampler", "ksc_tau", "ksc_mu", "ksc_pi"),
  test_fns = test_fns,
  parallelize = TRUE
)
```

```{r}
SBC::plot_rank_hist(primiceri_sbc_wrong$sbc_result)
SBC::plot_ecdf_diff(primiceri_sbc_wrong$sbc_result)
```

SBC checks were a lot clearer and more informative for this problem. There is a clear difference between the histograms & ECDF plots of the correct sampler and the wrong assembly one, flagging the error. The only variable that defies this is `s`, potentially due to being a discrete variable taking on only values {1, ..., 7}, leading to higher variance and noise.

